() ["./put.png"]

```{r, echo=FALSE, cache=TRUE}
library("plotly")
library("dplyr")
library("tidyr")
library("stringr")
library("ggplot2")
```

```{r, echo=FALSE}
education_order <- c(
    "I never completed any formal education",
    "Something else",
    "Primary/elementary school",
    "Secondary school",
    "Associate degree",
    "Some college/university study without earning a degree",
    "Bachelor’s degree",
    "Master’s degree",
    "Professional degree"
)

res2018raw <- read.csv("./developer_surveys/survey_results_2018.csv", sep = ",", header = TRUE)
res2018 <- res2018raw %>%
    select(FormalEducation, DevType, Salary = ConvertedSalary, SalaryType, Experience = YearsCodingProf, AdmiredLanguage = LanguageDesireNextYear, CompanySize) %>%
    mutate(Salary = as.numeric(Salary)) %>%
    select(-SalaryType) %>%
    filter(!is.na(Salary) & Salary != 0) %>%
    mutate(Education = str_trim(str_extract(FormalEducation, "^[^(]+"))) %>%
    mutate(Education = ifelse(Education == "Other doctoral degree", "Professional degree", Education)) %>%
    mutate(Education = factor(Education, ordered = TRUE, levels = education_order))


res2023raw <- read.csv("./developer_surveys/survey_results_2023.csv", sep = ",", header = TRUE)
res2023 <- res2023raw %>%
    select(EdLevel, DevType, Salary = ConvertedCompYearly, Experience = YearsCodePro, CompanySize = OrgSize, Country) %>%
    mutate(Education = str_trim(str_extract(EdLevel, "^[^(]+"))) %>%
    mutate(Education = factor(Education, ordered = TRUE, levels = education_order))
```
# How income is distributed based on education and type


```{r}
# education_devtype_salary <- res2018 %>%
#     filter(!is.na(DevType) & !is.na(Education)) %>%
#     filter(Education != "I never completed any formal education") %>%
#     separate_rows(DevType, sep = ";") %>%
#     group_by(Education, DevType) %>%
#     summarize(MedSalary = median(Salary))


# g <- ggplot(data = education_devtype_salary, aes(x = Education, y = MedSalary, group = DevType, color = DevType)) +
#     geom_line() +
#     scale_fill_brewer(palette = "Blues")
# g

# ggplotly(g)
```


# How experience vs salary
## Bar (x - experience) y(salary)

```{r, cache=TRUE, echo=FALSE}
experience_levels <- c(
    "0-2", "3-5", "6-8", "9-11",
    "12-14", "15-17", "18-20", "21-23",
    "24-26", "27-29", "30 or more"
)

discretize_experience <- function(experience) {
    experience <- as.numeric(experience)
    return(case_when(
        experience >= 0 & experience <= 2 ~ "0-2",
        experience >= 3 & experience <= 5 ~ "3-5",
        experience >= 6 & experience <= 8 ~ "6-8",
        experience >= 9 & experience <= 11 ~ "9-11",
        experience >= 12 & experience <= 14 ~ "12-14",
        experience >= 15 & experience <= 17 ~ "15-17",
        experience >= 18 & experience <= 20 ~ "18-20",
        experience >= 21 & experience <= 23 ~ "21-23",
        experience >= 24 & experience <= 26 ~ "24-26",
        experience >= 27 & experience <= 29 ~ "27-29",
        TRUE ~ "30 or more"
    ))
}

exp_sal18 <- res2018 %>%
    select(Experience, Salary) %>%
    na.omit() %>%
    mutate(Experience = gsub(" years", "", Experience)) %>%
    mutate(Experience = factor(Experience, levels = experience_levels, ordered = TRUE)) %>%
    group_by(Experience) %>%
    summarize(MedSalary = median(Salary)) %>%
    mutate(year = "2018")

exp_sal23 <- res2023 %>%
    select(Experience, Salary) %>%
    na.omit() %>%
    mutate(Experience = discretize_experience(Experience)) %>%
    mutate(Experience = factor(Experience, levels = experience_levels, ordered = TRUE)) %>%
    group_by(Experience) %>%
    summarize(MedSalary = median(Salary)) %>%
    mutate(year = "2023")

exp_sal <- union(exp_sal18, exp_sal23)
```

```{r}
ggplot(exp_sal, aes(x = Experience, y = MedSalary, fill = year, color = year)) +
    geom_bar(stat = "identity", position = position_dodge(), width = 0.45)
# facet_wrap(~year)
```


## Inflation plot
```{r, echo=FALSE}
usd_inflation <- data.frame(
    Year = c(2018, 2019, 2020, 2021, 2022, 2023),
    Inflation = c(2.44, 1.81, 1.23, 4.70, 8.00, 4.10)
)

g <- ggplot(usd_inflation, aes(x = Year, y = Inflation)) +
    geom_smooth() +
    geom_point()

total_inflation <- prod(1 + usd_inflation$Inflation / 100)

salary_increase <- (exp_sal23$MedSalary - exp_sal18$MedSalary) / exp_sal18$MedSalary

salary_increase_df <- data.frame(
    Experience = exp_sal23$Experience,
    SalaryIncrease = salary_increase
)

g <- ggplot(salary_increase_df, aes(x = Experience, y = SalaryIncrease)) +
    geom_bar(stat = "identity") +
    geom_hline(yintercept = total_inflation - 1, color = "black", linetype = "dashed") +
    annotate("text", x = 8, y = total_inflation - 0.97, label = "Total inflation between 2018 and 2023", color = "black", size = 4) +
    scale_y_continuous(labels = scales::percent, limit = c(-0.5, 0.5), breaks = sort(c(seq(-0.5, 0.5, 0.1), total_inflation - 1)))
ggplotly(g)
```

## With inflation

# Are admired languages well paid
```{r}
admired18 <- res2018 %>%
    select(AdmiredLanguage, Salary) %>%
    separate_rows(AdmiredLanguage, sep = ";") %>%
    group_by(AdmiredLanguage) %>%
    summarize(SalaryMed = median(Salary), AdmiredLanguageCount = n())

g <- ggplot(admired18, aes(x = AdmiredLanguageCount, y = SalaryMed, color = AdmiredLanguage, fill = AdmiredLanguage)) +
    geom_jitter(show.legend = FALSE) +
    geom_text(
        aes(label = AdmiredLanguage),
        vjust = case_when(
            admired18$AdmiredLanguage == "Python" ~ 1.5,
            admired18$AdmiredLanguage == "CSS" ~ 1.8,
            admired18$AdmiredLanguage == "Visual Basic 6" ~ 1,
            admired18$AdmiredLanguage == "C++" ~ 1,
            admired18$AdmiredLanguage == "Perl" ~ 1,
            TRUE ~ -0.5
        ),
        hjust = -0.2,
        show.legend = FALSE
    ) +
    theme(legend.position = "none")

ggplotly(g)
```


## Admired salary - language

# Is it better to work in big companies?
## Subplots x age_bin, and comparison of salaries per company size

```{r}
company_levels <- c(
    "Freelancer",
    "2 to 9 employees",
    "Fewer than 10 employees",
    "10 to 19 employees",
    "20 to 99 employees",
    "100 to 499 employees",
    "500 to 999 employees",
    "1,000 to 4,999 employees",
    "5,000 to 9,999 employees",
    "10,000 or more employees"
)

company_salary18 <- res2018 %>%
    select(
        CompanySize, Salary, Experience
    ) %>%
    na.omit() %>%
    mutate(Experience = gsub(" years", "", Experience)) %>%
    mutate(Experience = factor(Experience, levels = experience_levels, ordered = TRUE)) %>%
    mutate(CompanySize = factor(CompanySize, levels = company_levels, ordered = TRUE)) %>%
    group_by(CompanySize, Experience) %>%
    # summarize(SalaryMed = format(median(Salary), scientific=FALSE))
    summarize(SalaryMed = median(Salary))

company_salary23 <- res2023 %>%
    select(
        CompanySize, Salary, Experience
    ) %>%
    na.omit() %>%
    filter(CompanySize != "I don’t know") %>%
    mutate(CompanySize = ifelse(CompanySize == "Just me - I am a freelancer, sole proprietor, etc.", "Freelancer", CompanySize)) %>%
    mutate(Experience = discretize_experience(Experience)) %>%
    mutate(Experience = gsub(" years", "", Experience)) %>%
    mutate(Experience = factor(Experience, levels = experience_levels, ordered = TRUE)) %>%
    mutate(CompanySize = factor(CompanySize, levels = company_levels, ordered = TRUE)) %>%
    group_by(CompanySize, Experience) %>%
    summarize(SalaryMed = median(Salary))



ggplot(company_salary18, aes(x = CompanySize, y = SalaryMed, fill = CompanySize), format(y, scientific = FALSE)) +
    geom_bar(stat = "identity") +
    facet_wrap(~Experience) +
    scale_y_continuous(labels = scales::comma, breaks = seq(0, 200000, 25000)) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )

ggplot(company_salary23, aes(x = CompanySize, y = SalaryMed, fill = CompanySize), format(y, scientific = FALSE)) +
    geom_bar(stat = "identity") +
    facet_wrap(~Experience) +
    scale_y_continuous(labels = scales::comma, breaks = seq(0, 200000, 25000)) +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
    )
```

## Are differences statistically significant?

# Wher to work?
## Map of the world with colored salaries

# Best paying countries

Of course salary depends greatly on the country you work with. 

```{r world_map, echo=FALSE}
convert_country <- function(country) {
    return(case_when(
        country == "Antigua and Barbuda" ~ "Antigua",
        country == "Brunei Darussalam" ~ "Brunei",
        country == "Côte d'Ivoire" ~ "Ivory Coast",
        country == "Iran, Islamic Republic of..." ~ "Iran",
        country == "Lao People's Democratic Republic" ~ "Laos",
        country == "Libyan Arab Jamahiriya" ~ "Libya",
        country == "Republic of Moldova" ~ "Moldova",
        country == "Russian Federation" ~ "Russia",
        country == "Syrian Arab Republic" ~ "Syria",
        country == "The former Yugoslav Republic of Macedonia" ~ "North Macedonia",
        country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
        country == "United Republic of Tanzania" ~ "Tanzania",
        country == "United States of America" ~ "USA",
        country == "Venezuela, Bolivarian Republic of ..." ~ "Venezuela",
        country == "Viet Nam" ~ "Vietnam",
        TRUE ~ country
    ))
}

map_data23 <- res2023 %>%
    select(Country, Salary) %>%
    mutate(Country = convert_country(Country)) %>%
    na.omit() %>%
    group_by(Country) %>%
    summarize(SalaryMed = median(Salary))

WorldData <- map_data("world") %>%
    filter(region != "Antarctica") %>%
    fortify()

p <- ggplot() +
    geom_map(
        data = WorldData, map = WorldData,
        aes(x = long, y = lat, group = group, map_id = region),
        fill = "white", colour = "#7f7f7f", size = 0.5
    ) +
    geom_map(
        data = map_data23, map = WorldData,
        aes(fill = SalaryMed, map_id = Country),
        colour = "#7f7f7f", size = 0.5
    ) +
    coord_map("rectangular", lat0 = 0, xlim = c(-180, 180), ylim = c(-60, 90)) +
    scale_fill_continuous(low = "thistle2", high = "darkred", guide = "colorbar") +
    scale_y_continuous(breaks = c()) +
    scale_x_continuous(breaks = c()) +
    labs(fill = "legend", title = "Title", x = "", y = "") +
    theme_bw()
p
```
As shown above, the countries which stand out are USA, switzerland and Australia

